---
title: "Class 8"
author: "Tavleen Takhar"
date: "10/25/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-mean Clustering 

Our first example with **kmeans()** function

```{r}
# Example plot to see how Rmarkdown works
plot(1:10, typ="l")
```

Back to Kmeans...

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))

plot(x)
```


```{r}
k <- kmeans(x, center=2, nstart=20)
k
```

# How many points are in each cluster?
```{r}
k$size
```
# What component of your result object details for assignment/membership?
```{r}
k$cluster
```
# What component of your result objects details for cluster center?
```{r}
k$centers
```

```{r}
plot(x, col=k$cluster)
palette(c("blue", "red"))
plot(x, col=k$cluster)
points(k$centers, col="green", pch=20, cex=3)
```
```{r}
k <- kmeans(x, center=2, nstart=20)
k3 <- kmeans(x, center=3, nstart=20)
k$tot.withinss
k3$tot.withinss
```












# Hierarchial Clustering in R

```{r}
d <- dist(x)
hc <- hclust(d)
plot(hc)
```


```{r}
plot(hc)
abline(h=8, col="red")

cutree(hc, h=8)
```




### A more "real" example of data clustering

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col, pch=20)
```




```{r}
d <- dist(x)
hc <- hclust(d)
plot(hc)

```
```{r}
grp.2 <- cutree(hc, k=2)
grp.3 <- cutree(hc, k=3)
plot(hc)

```



```{r}
table(grp.3, col)
```



# Principal Component Analysis
```{r}
## Initialize a blank 100 row by 10 column matrix
mydata <- matrix(nrow=100, ncol=10)
## Lets label the rows gene1, gene2 etc. to gene100
rownames(mydata) <- paste("gene", 1:100, sep="")
## Lets label the first 5 columns wt1, wt2, wt3, wt4 and wt5
## and the last 5 ko1, ko2 etc. to ko5 (for "knock-out")
colnames(mydata) <- c( paste("wt", 1:5, sep=""),
 paste("ko", 1:5, sep="") )
## Fill in some fake read counts
for(i in 1:nrow(mydata)) {
 wt.values <- rpois(5, lambda=sample(x=10:1000, size=1))
 ko.values <- rpois(5, lambda=sample(x=10:1000, size=1))

 mydata[i,] <- c(wt.values, ko.values)
}
head(mydata)
```

# Now lets try to find structure in this dat
```{r}
head(t(mydata))

```

```{r}
pca <- prcomp(t(mydata), scale=TRUE)
pca
```

# Make a PC plot of PC1 vs PC2. For this I use the `$x` component of our output 
```{r}
plot(pca$x[,1], pca$x[,2])
```




